The Hydrogen Epoch of Reionization Array Dish I: Beam Pattern Measurements and Science Implications

### 2016-02-12 DishBeamMeasurements/paper/DishBeamMeasurementsPaper_rev4.tex

\author{Abraham R. Neben\altaffilmark{1},
Richard F. Bradley\altaffilmark{2,3},
Jacqueline N. Hewitt\altaffilmark{1},
David R. DeBoer\altaffilmark{4},
Aaron Parsons\altaffilmark{4},
James E. Aguirre\altaffilmark{6},
Aaron Ewall-Wice\altaffilmark{1},
Nipanjana Patra\altaffilmark{4},
Nithyanandan Thyagarajan\altaffilmark{5},
Zaki S. Ali\altaffilmark{4},
Judd Bowman\altaffilmark{5},
Carina Cheng\altaffilmark{4},
Roger Dickenson\altaffilmark{3},
Phillip Doolittle\altaffilmark{3},
Dennis Egan\altaffilmark{3},
Mike Hedrick\altaffilmark{3},
Saul A. Kohn\altaffilmark{6},
Benjamin R.B. Saliwanchik\altaffilmark{7},
Patrick Schaffner\altaffilmark{3},
John Shelton\altaffilmark{3},
H.A. Taylor\altaffilmark{3},
Rusty Taylor\altaffilmark{3},
Butch Wirt\altaffilmark{3},
Jeff Zheng\altaffilmark{1}}

\affil{\altaffilmark{1}MIT Kavli Institute, Massachusetts Institute of Technology, Cambridge, MA, 02139 USA}
\affil{\altaffilmark{2}Dept. of Electrical and Computer Engineering, University of Virginia, Charlottesville, VA 22904}
\affil{\altaffilmark{3}National Radio Astronomy Obs., Charlottesville, VA}
\affil{\altaffilmark{4}Dept. of Astronomy, University of California, Berkeley, CA, USA}
\affil{\altaffilmark{5}Arizona State University, School of Earth and Space Exploration, Tempe, AZ 85287, USA}
\affil{\altaffilmark{6}Dept. of Physics and Astronomy, University of Pennsylvania, Philadelphia, PA}
\affil{\altaffilmark{7}Astrophysics and Cosmology Research Unit, University of KwaZulu-Natal, Durban, South Africa}

@@

\begin{abstract}
The Hydrogen Epoch of Reionization Array (HERA) is a radio interferometer aiming to detect the power spectrum of 21\,cm fluctuations from neutral hydrogen from the Epoch of Reionization. Drawing on lessons from the Murchison Widefield Array (MWA) and the Precision Array for Probing the Epoch of Reionization (PAPER), HERA is a hexagonal array of large (14\,m diameter) dishes with suspended dipole feeds with element collecting area of 100\,m$^2$. Not only does the dish determine overall sensitivity, it determines the aggregate frequency structure of foregrounds in the interferometer. This is the first of a series of four papers characterizing the frequency and angular response of the with simulations and measurements. We focus in this paper on the angular response (i.e., power pattern), which sets the relative weighting between sky regions of high and low delay, and thus, apparent source frequency structure. We measure the angular response at 137\,MHz using the ORBCOMM beam mapping system of \citet{neben15}. We measure a collecting area of 93\,m$^2$ in the optimal dish/feed configuration, implying HERA-331 should detect the EOR power spectrum with a signal-to-noise ratio of 19.3 using a foreground avoidance approach, and 74.3 using a foreground subtraction approach. Lastly we study the impact of these beam measurements on the distribution of foregrounds in Fourier space.
\end{abstract}

\keywords{instrumentation: interferometers --- techniques: interferometric --- cosmology: observations --- dark ages, reionization, first stars}

@@

A new generation of low frequency radio telescopes is coming online with the goal of
 probing redshifted 21\,cm emission from the Cosmic Dawn. These observations will 
 complement indirect probes of the Dark Ages and Epoch of Reionization such as quasar 
 sightlines and the CMB optical depth which leave the reionization 
 history of the universe only loosely constrained. (See \citet{FurlanettoReview, miguelreview, PritchardLoebReview, aviBook, zaroubi} for reviews) Sensitivity and foreground removal are 
 the main challenges in 21\,cm observations, as the expected cosmological signal is 4--5 
 orders of magnitude fainter in brightness temperature than Galactic and extragalactic foregrounds. Radio 
 interferometers such as the Murchison Widefield Arry (MWA) \citep{tingay13,mwascience}, the Precision Array for Probing the Epoch of Reionization (PAPER) \citep{ali2015}, the Giant Meterwave Radio Telescope (GMRT) 
 \citep{Paciga2011}, and the Low Frequency Array (LOFAR) \citep{lofar} are seeking a first detection of 
 cosmological 21\,cm emission in power spectrum measurements, where the smooth 
 frequency evolution of the foreground emission separates from the spectrally 
 rough cosmological signal whose frequency dimension probes a line of sight through the 
 inhomogenous reionizing universe.

The Hydrogen Epoch of Reionization Array (HERA) \citep[][deBoer et al. (in prep)]{PoberNextGen} is drawing on lessons learned by the MWA and PAPER to reach the calibration and foreground isolation accuracy required to make a significant detection and characterization of the cosmological signal. HERA uses 14\,m diameter parabolic dishes arranged in a compact, hexagonal array to achieve coherent integration of the very low surface brightness 21\,cm signal. Redundant baselines also permit redundant calibration techniques which solve for the relative calibration between all antennas \citep{liu2010,zheng14}. A central lesson from first generation instruments is that it is essential to characterize the instrument response to foreground emission lest instrument frequency dependence smear foreground power into cosmological signal modes.

@@

In an ideal achromatic instrument the foreground emission would be confined to the lowest few 
line of sight Fourier modes \citep[e.g.,][]{MoralesBowmanHewittFGsub}, however  the  
interferometer's frequency-dependent point spread function smears foreground power into a ``wedge'' shaped 
region in  $(k_\perp,k_\parallel)$ Fourier space \citep{Dattapowerspec,X13, PoberWedge,MoralesPSShapes, VedanthamWedge, nithya13, CathWedge, AdrianWedge1, AdrianWedge2,parsons12b}. This effect is straightforward to understand for a single baseline which measures the sky intensity weighted by the complex sky fringe 
$e^{i \vec{k}\cdot\vec{b}}$, where $\vec{k}(\nu)$ is the wave vector of incident radiation pointing along its direction of travel, $\vec{b}$ is the baseline vector in meters, and $\nu$ is the observation frequency. Thus sources at different positions relative to the baseline vector manifest different 
frequency structure despite their intrinsically smooth spectra, but are geometrically limited 
by the baseline length to a maximum frequency dependence of  $e^{2\pi i \nu \tau_g}$, where $\tau_g=\vec{b}\cdot\hat{k}/c$ is the delay in radiation arrival time between the two antennas in the baseline. This limits the foreground contamination 
to a wedge shaped region in Fourier space with $k_\parallel<a k_\perp$, where $k_\perp$ and $k_\parallel$ represent spatial modes 
perpendicular and parallel to the line of sight, and $a$ is a constant depending on the observational frequency and cosmology. The complement of the wedge is known as the ``EOR window''.

It is convenient to phrase this description in terms 
of the delay in radiation arrival at the baseline's two antennas, $\tau$, where $\tau_
\text{max}=b/c$. Sources at low delay have little frequency 
structure, while those near $\tau=\tau_\text{max}$ acquire the maximum frequency 
structure given the baseline length. The high delay regions of the sky lie near the horizon while low delay 
regions lie closer to zenith and also perpendicular to the baseline vector.

@@

For foreground avoidance-based power spectrum estimation, so long as foreground emission is perfectly contained in the wedge it is irrelevant how much or 
little of it there is, but real world effects smear power beyond the geometrical edge of the wedge into the EOR 
window. Finite bandwidth, imperfect bandpass calibration, and faraday rotation \citep{jelic2010,moore2013,giannisurvey, moore2013,moore2015,asad2015} of polarized sources can all imprint slight frequency structure on otherwise spectrally smooth sources, and those appearing closest to the edge of the wedge are 
most at risk of leaking into the EOR window. In fact, 
\citet{nithya15,nithya15b} observe in simulations and then in data that while naively we might expect minimal emission 
at the very edge of the wedge because typical near-horizon beam responses are so small, 
two effects can cause a relative brightening of emission at those maximal delays, creating a characteristic ``pitchfork'' shape. This horizon 
brightening is caused by the large solid angle subtended by the near-horizon regions of the 
sky, as well as the apparent shortening of baselines when viewed nearly on axis at these elevations. 
This second effect makes intermediate length baselines of tens to hundreds of meters sensitive to the very bright diffuse 
emission they would not see from near zenith. Together, these effects can overcome the decline in beam sensitivity near the horizon. 
All these considerations highlight the antenna beam as a critical design parameter for 
21\,cm observatories.

@@

This is the first in a series of four papers detailing the HERA element. In this work we study 
the angular response of the dish and its implications for power spectrum measurements. The three companion 
papers present reflectometry measurements (Patra et al., submitted) and simulations (Ewall-Wice et al., 
submitted) of the dish frequency response, as well as detailed 
foreground simulations for HERA (Thyagarajan et al., submitted). A general description to the design of the 
HERA experiment is given by DeBoer et al. (in prep). In essence, we 
require a large collecting area for
 sensitivity and minimal sidelobes and horizon response without incurring the large cost per collecting 
area of very large dishes. A dish is preferred to a large phased array as it has a less complex beampattern and reduced potential for antenna-to-antenna variation (Neben et al, submitted). The 331 dishes are positioned in a compact, hexagonal 
array permitting redundant baseline calibration and coherent integration in $
\vec{k}$ space \citep{zheng14,parsons12a}. 
%The smaller field of view of the HERA dish relative to that of the PAPER dipole or the MWA tile does not limit sensitivity in Fourier space as most leverage on high $k$ modes is from line of sight modes anyway \citep{PoberNextGen}.

@@

In this paper we first characterize the angular response of a prototype HERA dish at the National Radio 
Astronomy Observatory--Green Bank. We use the beam mapping system of \citet{neben15} to 
measure the 137\,MHz beam pattern using the ORBCOMM satellite constellation. We obtain beam 
measurements out to zenith angles of $\sim60^\circ$ where the beam response is -35\,dB relative to zenith, and compare with different numerical electromagnetic models. We characterize the dish beam at various feed heights to map out the focus and study beam errors due to feed misalignment. We compute the collecting areas and implied EOR power spectrum sensitivities of our measured beams. After verifying our models, we consider the science implications of these 
beam patterns by foreground delay spectra at different baseline lengths and observing conditions to study when the horizon brightening effect is strongest, and thus, when foreground are most at risk of leaking into the EOR window.

@@

The HERA element (Fig. \ref{fig:feeddiagram}) is a 14\,m diameter faceted parabolic dish ($f\#=0.32$) with a dual-polarized dipole feed suspended at prime focus. The dish surface is formed by wire mesh sheets (i.e., facets) mounted on PVC tubes which run from the lip of the dish to the hub at the vertex. The feed consists of a dual linear polarization PAPER sleeved dipole mounted 17'' below a 78'' diameter wire mesh back plane surrounded by a 30'' deep cylinder, suspended from cables attached to each the three telephone poles around the dish. The dipole ``sleeves'' are circular disks just above and below the dipole designed to broaden its frequency response. The cylinder is offset 0.5'' from the back plane, and is designed to make the beams for both linear polarizations more similar, and also to taper the dipole beam towards the edge of the dish to mitigate dish-dish coupling. Fig. \ref{fig:feedphoto} shows the feed as deployed on the ground for early testing.  The nominal dish focus is $f=(f\#)D=4.48$\,m, though our numerical electromagnetic models suggest the actual focus is slightly higher due to the feed geometry. In this work we study the dish beam pattern at feed heights of 4.5\,m, 5.0\,m, and 5.3\,m, measured from dish surface to feed plane, the last height being the maximum height we can achieve with the feed suspension system installed on the dish. These height measurements are uncertain at the $\pm5\%$ level in this study. For more details on the dish design and construction see DeBoer et al. (in prep). Feed/dish optimization studies are ongoing and the values of these parameters may change in the full HERA array.

\begin{figure}[h]
\includegraphics[width=3.4in]{dish_and_feed_diagram.pdf}
\caption{Diagram showing the dimensions and layout of the HERA dish and suspended feed.}
\label{fig:feeddiagram}
\end{figure}

\begin{figure}[h]
\includegraphics[width=3.4in]{feed.jpg}
\caption{Prototype HERA feed set up outside the dish for preliminary characterization. This feed revision consists of a dual-polarized sleeved dipole offset 17'' from a 78'' diameter back plane, surrounded by a 30'' deep cylindrical skirt.}
\label{fig:feedphoto}
\end{figure}

As the HERA element is larger than the MWA or PAPER antenna elements, one might worry about the  smaller field of view and thus smaller number of measured Fourier modes perpendicular to the line of sight. However, this is a 
small effect for 21\,cm power spectrum analyses as our leverage on $k$ modes comes primarily from modes along the line of sight (in the frequency dimension). Further, HERA's smaller field of view is actually desirable in that it drastically reduces the magnitude of emission at the edge of the wedge compared to a simple dipole element \citep{nithya15}. A second potential drawback is frequency structure introduced by time domain reflections between the dish and feed detailed by Ewall-Wice et al. (submitted) with simulations and Patra et al. (in prep) with zenith reflectometry measurements. These works demonstrate, though, that the slight frequency structure of the dish is sufficiently small to not interfere with EOR science.
\\
\subsection{Dish Modeling}
\label{sec:dishmodels}

\begin{figure*}
\centering
\includegraphics[width=7in]{dave197_rich195_airy_beams.pdf}
\caption{Simulated dish power patterns (NS polarization) at 137\,MHz (see Sec. \ref{sec:dishmodels}) with $h_\text{feed}=5$\,m using HFSS (left) and CST (middle) are shown beside an ideal Airy pattern for a 14\,m diameter dish for comparison.}
\label{fig:modelbeams}
\end{figure*}

We numerically model the HERA dish in both ANSYS HFSS\footnote{http://www.ansys.com/Products/Electronics/ANSYS-HFSS} and in CST Microwave Studio\footnote{https://www.cst.com/Products/CSTMWS}, making slightly different assumptions, in order to study the dependence of the beam pattern on modeling inaccuracies and material uncertainties. In particular, the near horizon beam response, which sets the level of horizon brightening in the delay spectrum, is quite sensitive to modeling assumptions. Modeling also probes the best focus of the dish--feed system.

The HFSS model assumes a solid 14\,m parabolic dish ($f\#=0.32$) with a 1\,m hole at the vertex. In reality a 1\,m diameter circle at the vertex is filled with concrete and supports the PVC pipes which suspend the mesh panels, but we effectively assume here it is perfectly absorbant.  The feed is modeled as sleeved dipole with a solid metal back plane and cylinder with the dimensions given above. Copper is used for the dipole itself, and aluminum for all reflecting surfaces. In the simulations, we excite one of the dipoles using a modal port and measure the total gain response in each direction to 137.5\,MHz radiation.

The CST model again assumes a 14\,m parabolic diameter solid aluminum paraboloid with $f\#=0.32$ but neglects the 1\,m hole at the vertex, effectively assuming the concrete and earth ground behind it are perfectly reflective. The actual material properties are somewhere between fully reflective and fully absorbant, and thus, between these two models. The feed model has the same dimensions and properties as in the HFSS model. These simulations are done by exciting one of the dipoles with band-limited noise (100--200\,MHz), then measuring the farfield radiation in each direction after the excitation pulse energy within the structure decays to below -80\,dB.

The simulated HFSS and CST beams for the NS dipole are plotted in Fig. \ref{fig:modelbeams} (left and center panel) along with an Airy pattern for comparison. As expected, both model beams have slightly stronger sidelobes and wider main lobes than the ideal Airy pattern. The dipole sleeve (circular pieces in Fig. \ref{fig:feedphoto}) and skirt result in a feed beam which is slightly elongated in the E plane and slightly compressed in the H plane, opposite to the behavior of a simple dipole. This wider dish illumination in the NS direction by the NS feed dipole results in a narrower dish beam in the NS direction. Similarly, the EW dish beam is narrower in the EW direction. Lastly, we note that in both models, the best focus is found to be close to 5.23\,m with this feed geometry, slightly larger than the nominal focus of 4.5\,m, as expected.

@@

Unlike celestial source beam measurements, where the flux may be 
assumed constant over the timescale of the measurement, satellite fluxes can vary rapidly 
due to changing distance, orientation, and transmission power. To correct for this, we 
measure the satellite flux in each ground polarization (EW and NS) using a simple, well-
modeled reference antenna. Comparison of this measured power with that observed in the 
Antenna-Under-Test (AUT) gives the AUT beam response in the direction of the satellite. 
An equivalent interpretation of the measurement is that the power ratio between the AUT and the reference 
antenna gives the relative beam response in the satellite direction, and multiplication by 
the reference antenna model yields the desired AUT response. As discussed in 
\citet{neben15}, this procedure correctly measures the response of the AUT to unpolarized radiation
 despite the fact that satellite signals are generally polarized.

In detail, we measure the dual-polarization RMS power received by each antenna in 512 2\,kHz 
channels across the 137--138\,MHz band. Each band power is averaged over $\sim0.2$
\,sec. There are 0--3 satellites above the horizon at any given time transmitting on different 
$\sim15$\,kHz wide sub-bands in 137--138\,MHz. By observing at many different 
frequencies, we probe the beam response in all these directions simultaneously. We 
compute the satellite positions using the orbital elements published by Celestrak
\footnote{http://www.celestrak.com/NORAD/elements/orbcomm.txt} and the orbital 
integrator \texttt{predict}\footnote{http://www.qsl.net/kd2bd/predict.html}. However, the 
satellite frequencies vary occasionally to avoid interference within the constellation. 
\citet{zheng14} use interferometric phases to identify and exclude times when multiple 
satellites are in view. As our data acquisition system makes only total power 
measurements, we instead use an ORBCOMM interface box (typically supplied to 
commercial users of the network) to connect to passing satellites and record their identifier 
and transmission frequency during each pass.

>>>>

\author{Abraham R. Neben\altaffilmark{1,2},
Richard F. Bradley\altaffilmark{3,4},
Jacqueline N. Hewitt\altaffilmark{1,2},
David R. DeBoer\altaffilmark{5},
Aaron R. Parsons\altaffilmark{5},
James E. Aguirre\altaffilmark{7},
Zaki S. Ali\altaffilmark{5},
Carina Cheng\altaffilmark{5},
Aaron Ewall-Wice\altaffilmark{1,2},
Nipanjana Patra\altaffilmark{5},
Nithyanandan Thyagarajan\altaffilmark{6},
Judd Bowman\altaffilmark{6},
Roger Dickenson\altaffilmark{4},
Joshua S. Dillon\altaffilmark{5},
Phillip Doolittle\altaffilmark{4},
Dennis Egan\altaffilmark{4},
Mike Hedrick\altaffilmark{4},
Daniel C. Jacobs\altaffilmark{6},
Saul A. Kohn\altaffilmark{7},
Patricia J. Klima\altaffilmark{4},
Kavilan Moodley\altaffilmark{8},
Benjamin R.B. Saliwanchik\altaffilmark{8},
Patrick Schaffner\altaffilmark{4},
John Shelton\altaffilmark{4},
H.A. Taylor\altaffilmark{4},
Rusty Taylor\altaffilmark{4},
Butch Wirt\altaffilmark{4},
Haoxuan Zheng\altaffilmark{1,2}}

\affil{\altaffilmark{1}MIT Kavli Institute, Massachusetts Institute of Technology, Cambridge, MA, 02139 USA}
\affil{\altaffilmark{2}Dept. of Physics, Massachusetts Institute of Technology, Cambridge, MA, 02139 USA}
\affil{\altaffilmark{3}Dept. of Electrical and Computer Engineering, University of Virginia, Charlottesville, VA 22904}
\affil{\altaffilmark{4}National Radio Astronomy Obs., Charlottesville, VA}
\affil{\altaffilmark{5}Dept. of Astronomy, University of California, Berkeley, CA, USA}
\affil{\altaffilmark{6}Arizona State University, School of Earth and Space Exploration, Tempe, AZ 85287, USA}
\affil{\altaffilmark{7}Dept. of Physics and Astronomy, University of Pennsylvania, Philadelphia, PA}
\affil{\altaffilmark{8}Astrophysics and Cosmology Research Unit, University of KwaZulu-Natal, Durban, South Africa}

@@

\begin{abstract}
The Hydrogen Epoch of Reionization Array (HERA) is a radio interferometer aiming to detect the power spectrum of 21\,cm fluctuations from neutral hydrogen from the Epoch of Reionization (EOR). Drawing on lessons from the Murchison Widefield Array (MWA) and the Precision Array for Probing the Epoch of Reionization (PAPER), HERA is a hexagonal array of large (14\,m diameter) dishes with suspended dipole feeds. Not only does the dish determine overall sensitivity, it affects the observed frequency structure of foregrounds in the interferometer. This is the first of a series of four papers characterizing the frequency and angular response of the dish with simulations and measurements. We focus in this paper on the angular response (i.e., power pattern), which sets the relative weighting between sky regions of high and low delay, and thus, apparent source frequency structure. We measure the angular response at 137\,MHz using the ORBCOMM beam mapping system of \citet{neben15}. We measure a collecting area of 93\,m$^2$ in the optimal dish/feed configuration, implying HERA-320 should detect the EOR power spectrum at $z\sim9$ with a signal-to-noise ratio of 19.3 using a foreground avoidance approach, and 74.3 using a foreground subtraction approach. Lastly we study the impact of these beam measurements on the distribution of foregrounds in Fourier space.
\end{abstract}

\keywords{instrumentation: interferometers --- cosmology: observations --- reionization, first stars}

@@

A new generation of low frequency radio telescopes is coming online with the goal of
 probing redshifted 21\,cm emission from the Cosmic Dawn. These observations will 
 complement indirect probes of the Epoch of Reionization such as quasar 
 sightlines and the CMB optical depth, which leave the reionization 
 history of the universe only loosely constrained. (See \citet{FurlanettoReview, miguelreview, PritchardLoebReview, aviBook, zaroubi} for reviews) In the longer term, 21\,cm observations are expected to improve constraints on cosmology \citep[e.g.,][]{mao08, liu15a,liu15b}. Sensitivity and foreground removal are 
 the main challenges in 21\,cm observations, as the expected cosmological signal is 4--5 
 orders of magnitude fainter in brightness temperature than Galactic and extragalactic foregrounds. Radio 
 interferometers such as the Murchison Widefield Array (MWA) \citep{lonsdale09,tingay13,mwascience}, the Precision Array for Probing the Epoch of Reionization (PAPER) \citep{parsons10,parsons14,ali2015}, the Giant Meterwave Radio Telescope (GMRT) 
 \citep{Paciga2011}, and the Low Frequency Array (LOFAR) \citep{lofar} are seeking a first detection of 
 cosmological 21\,cm emission in power spectrum measurements. In the power spectrum, the spectrally smooth foreground emission separates from the spectrally 
 rough cosmological signal whose frequency dimension probes a line of sight through the 
 inhomogenous reionizing universe.

The Hydrogen Epoch of Reionization Array (HERA) \citep{PoberNextGen,deboer16} is drawing on lessons learned by the MWA and PAPER to reach the calibration and foreground isolation accuracy required to make a significant detection and characterization of the cosmological signal. HERA uses 14\,m diameter parabolic dishes arranged in a compact, hexagonal array to achieve coherent integration of the very low surface brightness 21\,cm signal. Redundant baselines also permit redundant calibration techniques which solve for the relative calibration between all antennas \citep{redundant3, redundant4, liu2010,zheng14}. A central lesson from first generation instruments is that it is essential to characterize the instrument response to foreground emission lest instrument frequency dependence smear foreground power into cosmological signal modes.

@@

In an ideal achromatic instrument the foreground emission would be confined to the lowest few 
line of sight Fourier modes \citep[e.g.,][]{MoralesBowmanHewittFGsub}, however  the  
interferometer's frequency-dependent point spread function smears foreground power into a ``wedge'' shaped 
region in $(k_\perp,k_\parallel)$ Fourier space \citep{Dattapowerspec,X13, PoberWedge,MoralesPSShapes, VedanthamWedge, nithya13, CathWedge, AdrianWedge1, AdrianWedge2,parsons12b}, where $k_\parallel$ modes are along the line of sight and $k_\perp$ modes are perpendicular to it. This effect is straightforward to understand for a single baseline which measures the sky intensity weighted by the complex sky fringe 
$e^{2\pi i \nu \tau_g}$, where $\tau_g=\vec{b}\cdot\hat{s}/c$ is the delay in radiation arrival time at the second antenna relative to the first antenna of the baseline. Here $\nu$ is the observation frequency, $\vec{b}$ is the baseline vector, and $\hat{s}$ is the direction of the source. 
 Thus sources at different positions relative to the baseline vector appear with different 
frequency structure despite their intrinsically smooth spectra. However, this instrumental frequency structure is limited 
by the baseline length to a maximum frequency dependence of  $e^{2\pi i \nu b/c}$ for sources at maximum delay, near the horizon in line with the baseline vector. This limits the foreground contamination 
to a wedge shaped region in Fourier space with $k_\parallel<a k_\perp$, where $k_\perp$ and $k_\parallel$ represent spatial modes 
perpendicular and parallel to the line of sight, and $a$ is a constant depending on the observational frequency and cosmology. The complement of the wedge is known as the ``EOR window''.

@@

For foreground avoidance-based power spectrum estimation, so long as foreground emission is perfectly contained in the wedge it is irrelevant how much or 
little of it there is, but real world effects smear power beyond the geometrical edge of the wedge into the EOR 
window. Finite bandwidth, imperfect bandpass calibration, and faraday rotation of polarized sources can all imprint slight frequency structure on otherwise spectrally smooth sources \citep{jelic2010,giannisurvey, moore2013,moore2015,asad2015,newburgh14,shaw15}, and those closest to the edge of the wedge are 
most at risk of leaking into the EOR window. In fact, 
\citet{nithya15,nithya15b} observe in simulations and then in data that while naively we might expect minimal emission 
at the very edge of the wedge because typical near-horizon beam responses are so small, 
two effects can cause a relative brightening of emission at those maximal delays, creating a characteristic ``pitchfork'' shape. This horizon 
brightening is caused by the large solid angle subtended by the near-horizon regions of the 
sky, as well as the apparent shortening of baselines when viewed nearly on axis at these elevations. 
This second effect makes intermediate length baselines of tens to hundreds of meters sensitive to the very bright diffuse 
emission they would not see from near zenith. Together, these effects can overcome the decline in beam sensitivity near the horizon. 
All these considerations highlight the antenna beam as a critical design parameter for 
21\,cm observatories.

@@

This is the first in a series of four papers detailing the HERA element. In this work we study 
the angular response of the dish and its implications for power spectrum measurements. The three companion 
papers present reflectometry measurements \citep{patra16} and simulations \citep{ewallwice16} of the dish frequency response, as well as detailed 
foreground simulations for HERA \citep{nithya16}. A general description of the design of the 
HERA experiment is given by DeBoer et al. \citep{deboer16}. In essence, we 
require a large collecting area for
 sensitivity, and minimal sidelobes and horizon response without incurring the large cost per collecting 
area of very large dishes. A dish is preferred to a large phased array as it has a less complex beampattern and reduced potential for antenna-to-antenna variation \citep{neben16}. The core array consists of 320 dishes positioned on a compact, hexagonal 
grid \citep{dillonparsons16} permitting redundant baseline calibration and coherent integration in $
\vec{k}$ space \citep{zheng14,parsons12a}. Improved imaging is permitted by 30 outriggers, but these do not appreciably affect power spectrum sensitivity.
%The smaller field of view of the HERA dish relative to that of the PAPER dipole or the MWA tile does not limit sensitivity in Fourier space as most leverage on high $k$ modes is from line of sight modes anyway \citep{PoberNextGen}.

@@

In this paper we first characterize the angular response of a prototype HERA dish at the National Radio 
Astronomy Observatory--Green Bank. We use the beam mapping system of \citet{neben15} to 
measure the 137\,MHz beam pattern using the ORBCOMM satellite constellation. We obtain beam 
measurements out to zenith angles of $\sim60^\circ$ where the beam response is -35\,dB relative to zenith, and compare with different numerical electromagnetic models. We characterize the dish beam at various feed heights to map out the focus and study beam errors due to feed misalignment. We compute the collecting areas and implied EOR power spectrum sensitivities of our measured beams. After verifying our models, we consider the science implications of these 
beam patterns by foreground delay spectra at different baseline lengths and observing conditions to study when the horizon brightening effect is strongest, and thus, when foregrounds are most at risk of leaking into the EOR window.

@@

The HERA element (Fig. \ref{fig:feeddiagram}) is a 14\,m diameter faceted parabolic dish ($ f/D=0.32$) with a dual-polarized dipole feed suspended at prime focus \citep{dishmemo}. Here $f$ is the focal length of the dish and $D$ is the dish diameter. The dish surface is formed by wire mesh sheets (i.e., facets) mounted on PVC tubes which run from the lip of the dish to the hub at the vertex. For these tests, the feed consists of a dual linear polarization PAPER sleeved dipole mounted 17'' below a 78'' diameter wire mesh back plane surrounded by a 30'' deep cylinder, suspended from cables attached to each of the three telephone poles around the dish. The dipole ``sleeves'' are circular disks just above and below the dipole designed to broaden its frequency response. The cylinder is offset 0.5'' from the back plane, and is designed to make the beams for both linear polarizations more similar. Fig. \ref{fig:feedphoto} shows the feed as deployed on the ground for early testing.  The nominal dish focus is $f=(f/d)D=4.48$\,m, though given its faceted design, the dish does not have a single focus. Our numerical electromagnetic models suggest the best focus is slightly higher due to the feed geometry. In this work we study the dish beam pattern at rigging heights of 4.5\,m, 5.0\,m, and 5.3\,m, measured from dish surface to feed plane, the last height being the maximum height we can achieve with the feed suspension system installed on the dish. These height measurements are uncertain at the $\pm5\%$ level in this study. For more details on the dish design and construction see \citet{deboer16}. Feed/dish optimization studies are ongoing and the values of these parameters may change in the full HERA array \citep{feedoptimizationmemo}.

\begin{figure}[h]
\includegraphics[width=3.4in]{dish_and_feed_diagram.pdf}
\caption{Diagram showing the dimensions and layout of the parabolic HERA dish and suspended feed.}
\label{fig:feeddiagram}
\end{figure}

\begin{figure}[h]
\includegraphics[width=3.4in]{feed.jpg}
\caption{Prototype HERA feed seen here outside the dish and upside-down for preliminary characterization. This feed revision consists of a dual-polarized sleeved dipole offset 17'' from a 78'' diameter back plane, surrounded by a 30'' deep cylindrical skirt.}
\label{fig:feedphoto}
\end{figure}

As the HERA element is larger than the MWA or PAPER antenna elements, one might worry about the  smaller field of view and thus smaller range of Fourier space probed perpendicular to the line of sight. However, this is a 
small effect for 21\,cm power spectrum analyses as our leverage on $k$ modes comes primarily from modes along the line of sight (in the frequency dimension). Further, HERA's smaller field of view is actually desirable in that it drastically reduces the magnitude of emission at the edge of the wedge compared to a simple dipole element \citep{nithya15}. A second potential drawback is frequency structure introduced by time domain reflections between the dish and feed detailed by Ewall-Wice et al. \citep{ewallwice16} with simulations and \citep{patra16} with zenith reflectometry measurements. These works demonstrate, though, that the slight frequency structure of the dish is sufficiently small to not interfere with EOR science.

\subsection{Dish Modeling}
\label{sec:dishmodels}

\begin{figure*}
\centering
\includegraphics[width=7in]{dave195_rich195_airy_beams.pdf}
\caption{Simulated dish power patterns (NS polarization) at 137\,MHz (see Sec. \ref{sec:dishmodels}) with $h_\text{feed}=5$\,m using HFSS (left) and CST (middle) are shown beside an ideal Airy pattern for a 14\,m diameter dish for comparison. Dased lines mark zenith angles of 20$^\circ$, 40$^\circ$, 60$^\circ$, and 80$^\circ$.}
\label{fig:modelbeams}
\end{figure*}

We numerically model the HERA dish in both ANSYS HFSS\footnote{http://www.ansys.com/Products/Electronics/ANSYS-HFSS} and in CST Microwave Studio\footnote{https://www.cst.com/Products/CSTMWS}, making slightly different assumptions, in order to study the range of realistic beams given modeling inaccuracies and material imperfections. In particular, the near horizon beam response, which sets the level of horizon brightening in the delay spectrum, is quite sensitive to modeling assumptions.

The HFSS model uses a realistic faceted dish \citep{deboer16} ($f/D=0.32$) with a 1\,m diameter dielectric similar to dry soil at the vertex. In reality a 1\,m diameter circle at the vertex is filled with concrete and supports the PVC pipes which suspend the mesh panels, but we effectively assume here it is perfectly absorbant.  The feed is modeled as a sleeved dipole with a solid metal back plane and cylinder with the dimensions given above. Copper is used for the dipole itself, and aluminum for all reflecting surfaces. In the simulations, we excite one of the dipoles using a modal port and measure the total gain response in each direction to 137.5\,MHz radiation.

The CST model assumes a 14\,m diameter perfect paraboloid of solid aluminum with $f/D=0.32$ but neglects the 1\,m hole at the vertex, effectively assuming the concrete and earth ground behind it are perfectly reflective. The feed model has the same dimensions and properties as in the HFSS model. These simulations are done by exciting one of the dipoles with band-limited noise (100--200\,MHz), then measuring the farfield radiation in each direction after the excitation pulse energy within the structure decays to below -80\,dB.

The simulated HFSS and CST beams for the NS dipole are plotted in Fig. \ref{fig:modelbeams} (left and center panel) along with an Airy pattern for comparison. As expected, both model beams have slightly stronger sidelobes and wider main lobes than the ideal Airy pattern. The dipole sleeve (circular pieces in Fig. \ref{fig:feedphoto}) and skirt result in a feed beam which is slightly elongated in the E plane and slightly compressed in the H plane, opposite to the behavior of a simple dipole. This wider dish illumination in the NS direction by the NS feed dipole results in a narrower dish beam in the NS direction. Similarly, the EW dish beam is narrower in the EW direction. Lastly, we note that in both models, the best focus is found to be close to 5.23\,m with this feed geometry.

@@

Unlike celestial source beam measurements, where the flux may be 
assumed constant over the timescale of the measurement, satellite fluxes can vary rapidly 
due to changing distance, orientation, and transmission power. To correct for this, we 
measure the satellite flux in each ground polarization (East-West (EW) and North-South (NS)) using a simple, well-
modeled reference antenna. Comparison of this measured power with that observed in the 
Antenna-Under-Test (AUT) gives the AUT beam response in the direction of the satellite. 
An equivalent interpretation of the measurement is that the power ratio between the AUT and the reference 
antenna gives the relative beam response in the satellite direction, and multiplication by 
the reference antenna model yields the desired AUT response. As discussed in 
\citet{neben15}, this procedure correctly measures the desired response of the AUT to unpolarized radiation despite the fact that satellite signals are generally polarized.

In detail, we measure the dual-polarization RMS power received by each antenna in 512 2\,kHz 
channels across the 137--138\,MHz band. Each band power is averaged over $\sim0.2$
\,sec. There are 0--3 satellites above the horizon at any given time transmitting on different 
$\sim15$\,kHz wide sub-bands in 137--138\,MHz. By observing at many different 
frequencies, we probe the beam response in all these directions simultaneously. We 
compute the satellite positions using the orbital elements published by Celestrak\footnote{http://www.celestrak.com/NORAD/elements/orbcomm.txt} and the orbital 
integrator \texttt{predict}\footnote{http://www.qsl.net/kd2bd/predict.html}. However, the 
satellite frequencies vary occasionally to avoid interference within the constellation. 
\citet{zheng14} use interferometric phases to identify and exclude times when multiple 
satellites are in view. As our data acquisition system makes only total power 
measurements, we instead use an ORBCOMM interface box (typically supplied to 
commercial users of the network) to connect to passing satellites and record their identifier 
and transmission frequency during each pass.
